### 2025-03-09 ms.qmd
updates

```{r, model_sem_OB}
#| label: model_sem_OB
# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
# To simplify the random factors, and since there is no effect of experience with the prey in prey detection, latency was averaged across the two preys. Every other predictor added to the model was based on previous brms for each variable separated.
# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
#
source(here("R", "func.R"))
#
SEM_df <- clean_df %>%
  group_by(lizard_id, prey) %>%
  mutate(t_D = mean(t_D, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(prey == "Unknown")
#
#
data_sem_OB <- SEM_df %>%
  filter(region == "OB")
# Create the models
#
refit <- FALSE
#
#
if(refit){
  m_OB <- brm(
    bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch)) +
    bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch)) +
    bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)) +
    bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch)) +
  set_rescor(TRUE),
  family = gaussian(),
  data = data_sem_OB, 
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OB, file = here("output/m_SEM/m_OB.rds"))
} else {
  m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
}
```

>>>>

```{r, model_sem_OB}
#| label: model_sem_OB
# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
# To simplify the random factors, and since there is no effect of experience with the prey in prey detection, latency was averaged across the two preys. Every other predictor added to the model was based on previous brms for each variable separated.
# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
#
source(here("R", "func.R"))
#
SEM_df <- clean_df %>%
  group_by(lizard_id, prey) %>%
  mutate(t_D = mean(t_D, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(prey == "Unknown")
#
#
data_sem_OB <- SEM_df %>%
  filter(region == "OB") %>%
  mutate(obs = as.integer(c(1:80))) %>%
  mutate(vec = rep(0, length(obs)))
# Create the models
#
refit <- FALSE
#
#
if(refit){
  m_OB <- brm(
    bf(t_D | se(vec, sigma = TRUE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
    bf(mean_dnadamage | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
    bf(mean_peroxidation | se(vec, sigma = TRUE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
    bf(mean_ros | se(vec, sigma = TRUE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
  set_rescor(FALSE),
  family = gaussian(),
  data = data_sem_OB,
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OB, file = here("output/m_SEM/m_OB.rds"))
} else {
  m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
}
#
#
#
#
#Chcking residual correlations
m_OB <- brm(
  bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
  bf(mean_dnadamage | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
  bf(mean_peroxidation | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
  bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
set_rescor(FALSE),
family = gaussian(),
data = data_sem_OB,
chains = 2, cores = 4, iter = 2000, warmup = 1000,
control = list(adapt_delta = 0.99, max_treedepth = 11))
#
m_OB2 <- brm(
  bf(t_D ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
  bf(mean_dnadamage ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
  bf(mean_peroxidation ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
  bf(mean_ros ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
set_rescor(FALSE),
family = gaussian(),
data = data_sem_OB,
chains = 2, cores = 4, iter = 2000, warmup = 1000,
prior = c(
    prior(normal(0, 1e-10), class = "sigma")),
control = list(adapt_delta = 0.99, max_treedepth = 11))
#
#
m_OB_rescorT <- readRDS(here("output/m_SEM/m_OB_rescorT.rds"))
```
